#Briefing 
Below you can find some useful functions for data manipulation
https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf 

```{r}
library(dplyr)
library(ggplot2)
library(readr)
```

```{r}
users = read_csv("/principal/materials/minor/midterm/users3.csv")
orders = read_csv("/principal/materials/minor/midterm/orders3.csv")
inventory = read_csv("/principal/materials/minor/midterm/inventory.csv")
```

## Descriptive statisitics

`dplyr::n()` Number of values in a vector   
`min()` Minimum value in a vector  
`max()` Maximum value in a vector  
`mean()` Mean value of a vector   
`median()` Median value of a vector  

## Logical expressions

**<** Less than   
**>** Greater than   
**!=** Not equal to   
**==** Equal to   
**is.na** Is NA   
**!is.na** Is not NA   
**<=** Less than or equal to   
**>=** Greater than or equal to   

**&** And  
`dplyr::filter(dataframe, logical_criteria & logical_criteria)`  
**|** Or  
`dplyr::filter(users, logical_criteria | logical_criteria)`

## Fast way to look at dataframes

`head()` - returns the first parts of vector or data frame  
`tail()` - returns the last parts of vector or data frame

## Data Manipulation

`filter(dataframe, logical_criteria)` Extract rows that meet logical criteria.
```{r}
dplyr::filter(users, country == "Russia")
```

`select(dataframe, column_names)` Select columns by name or helper function.
```{r}
dplyr::select(users, first_name, last_name, country)
```

`arrange(dataframe, column)` Order rows by values of a column (low to high).
```{r}
dplyr::arrange(users, first_name)
```

`top_n(dataframe, n, ordering_column)` Select the top n entries in each group, ordered by column.
```{r}
dplyr::top_n(inventory, 10, price)
```

`dplyr::left_join(dataframe_A, dataframe_B, by = "common_column")` # Join matching rows from b to a.
`dplyr::inner_join(dataframe_A, dataframe_B, by = "common_column")` # Join data. Retain only rows in both sets. 

```{r}
orders_with_description = dplyr::left_join(inventory, orders, by = "inventory_id")
```

`mutate(dataframe, new_column = expression)` Compute and append one or more new columns.
```{r}
dplyr::mutate(orders_with_description, sum_price = price * quantity)
```

### Summarize
`group_by(dataframe, column) %>% summarize(new_column = expression)` Compute separate summary row for each group. 
  
`group_by()` function breaks down a dataset into specified groups of rows. When you then apply `summarise()` on the resulting object they’ll be automatically applied “by group”. We use `summarise()` with aggregate functions, which take a vector of values and return a single number.

```{r}
dplyr::group_by(users, country) %>% dplyr::summarize(count = n())
```

### Tables


```{r, results='asis'}
country_summary = dplyr::group_by(users, country) %>% dplyr::summarize(count = n()) %>% arrange(desc(count))

pander::pandoc.table(top_n(country_summary, 5, count))
```

## Charts

### geom_histogram

`ggplot() +
  geom_histogram(data = ..., aes(x =continous variable), fill = to change the interior colouring , colour = to change the colour of bin outline,  binwidth = width of bins) + # The columns are positioned over a label that represents a quantitative variable
  ggtitle("...") + # title of the graph
  xlab("...") + # title of X axis
  ylab("...")` # title of Y axis
  
```{r}
ggplot() +
  geom_histogram(data = orders, aes(x = date), fill = "gray50", colour = "black")
```

### geom_bar

`ggplot() + 
  geom_bar(data = ..., aes(x =qualitative variable, y =..., fill = setting fill equal to a factor varible uses a discrete colour scale))` # The columns are positioned over a label that represents a qualitative (categorical) variable

```{r}
ggplot() + 
  geom_bar(data = users, aes(x = country, fill = gender))
```


### geom_boxplot

`ggplot() +
  geom_boxplot(data = ..., aes(x = qualitative variable, y = continous variable))` # good way to compare distributions for qualitative variables

```{r}
ggplot() +
  geom_boxplot(data = orders_with_description, aes(x = category, y = quantity)) 
```






